### 2021-09-17 paper/paper.tex
update paper

\documentclass[10pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{natbib}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage{subfigure}

@@

\begin{center}
\Huge
The Cover Tree Loss
\end{center}

@@

The main theoretical result for the cross entropy loss is that for a dataset with $n$ points,
the generalization error is $O(\sqrt{c/n})$.
There are many theorems that establish this result in different contexts.
The following is an example that relies on properties of SGD.
\begin{theorem}
\label{theorem:xentropy}
Assume that for all $i\in[n]$, $\ltwo{x_i} \le \rho$,
    and that for each class $i\in[c]$, $\ltwo{\w_i}\le B$.
Then the parameter vector $\bar W$ estimated from one epoch of SGD satisfies
\begin{equation}
    \E L_D(\bar W) - L_D(W^*) \le \frac {\sqrt cB\rho}{\sqrt n}
\end{equation}
where $W^* = \argmin L_D(W)$.
\end{theorem}
\begin{proof}
    \fixme{Write a proof of this result using Theorem 14.12 of Shalev-Shwartz and Ben-David and the properties of the cross entropy loss from the last set of notes.}
    \begin{align}
    {\| W \|^2_F}&=\sum_{i=1}^c{\| W_i\|^2}\\
                &\le\sum_{i=1}^c{B^2}\\
                &\le cB^2\\
   \| W \|_F  &\le \sqrt c B 
    \end{align}
Then we get:
    \begin{equation}
    l(\bar W)-l(W^*)\le\frac {\sqrt cB\rho}{\sqrt n}
    \end{equation}
   Take expectation of both side of equation(3):
   \begin{equation}
    \E L_D(\bar W) - L_D(W^*) \le \frac {\sqrt cB\rho}{\sqrt n}
\end{equation}
\end{proof}

@@

\fixme{Read the papers mentioned above.  Verify that you agree with the things I wrote.  Then, write a detailed mathematical description of the HSM loss and contrast it with the cover tree loss.}

\fixme{Read 5-10 additional papers about the hierarchical softmax.  For each paper, write a paragraph describing the paper's contributions and how it relates to the cover tree loss.  For the final paper we will reference these papers.  We will have to trim down the notes that you write for this task, but it's useful when writing the paper to have detailed notes about references.}

@@

\fixme{Write a detailed description of the sim loss}

Simloss\cite{Kobs2020SimLossCS} is similarity based loss function which incorporates class similarities to support the training of neural network classifiers.
Cross entropy loss assumes only one class is correct and punishes every misclassification, while Simloss indicates classes have similarities and classifying a target as a similar class is better.
When calculating the loss function, instead of only taking the probability vector at the target index Simloss also adds the product of similarity between target and other classes and probability vector at the corresponding class. 
Non zero values of similarity matrix lead to smaller losses when the network gives a high score to classes similar to the correct one. 
It would allow the classifier to make less severe mistakes as it learns to predict similar classes.
\begin{align}
    L_{Simloss}&=-\frac{1}{N}\log(\sum_{c=1}^{C}S_{y_i,c}\cdot p_i[c])\\
    % L_{Simloss}=-\frac{1}{N}\log(\sum_{c=1}^{C}S_{y_i,c}\cdot\frac {\exp(-\trans\w_{y_i} \x)}{\sum_{j=1}^C \exp(-\trans \w_j \x)})
    p_i&=\frac {\exp(-\trans\w_{y_i} \x)}{\sum_{j=1}^C \exp(-\trans \w_j \x)}
\end{align}

@@

This factorization decreases the statistical complexity of the problem from $O(\sqrt{c/n})$ to $O(\sqrt{e/n})$.
Recall that $e$ is a hyperparameter that is selected to be much less than $c$ and should represent the "intrinsic" dimensionality of the weight matrix.
The following theorem proves this result formally.
\begin{theorem}
Assume that for all $i\in[n]$, $\ltwo{x_i} \le \rho$,
    and that for each class $i\in[c]$, $\ltwo{\vv_i}\le 1$ and $\ltwo{\uu_i}\le B$.
Then the parameter vector $\bar W = V \bar U$ estimated from one epoch of SGD satisfies
\begin{equation}
    \E L_D(\bar W) - L_D(W^*) \le \frac {\sqrt eB\rho}{\sqrt n}
\end{equation}
where $W^* = \argmin L_D(W)$.
\end{theorem}
\begin{proof}
    \fixme{Write a proof of this result using Theorem 14.12 of Shalev-Shwartz and Ben-David.}
    \begin{align}
    {\| U \|^2_F}&=\sum_{i=1}^e{\| U_i\|^2}\\
                &\le\sum_{i=1}^e{B^2}\\
                &\le eB^2\\
   \| U \|_F  &\le \sqrt e B  \\
   \| W \|_F  &\le \sqrt e B
    \end{align}
    Then we get:
    \begin{equation}
    l(\bar W)-l(W^*)\le\frac {\sqrt eB\rho}{\sqrt n}
    \end{equation}
   Take expectation of both side of equation(8):
   \begin{equation}
    \E L_D(\bar W) - L_D(W^*) \le \frac {\sqrt eB\rho}{\sqrt n}
\end{equation}
\end{proof}

\section{The Tree Loss}

Like the weight matrix factorization, the tree loss works by reparameterizing the weight matrix for the softmax cross entropy loss.
The reparameterization is defined according to a \emph{label tree} structure that captures the similarity between labels.
Every class is represented by a leaf node in the label tree,
and internal nodes represent \emph{pseudo-classes} that group the existing classes into a logical hierarchy.
At this point, we assume that the label tree structure is given a priori, and in Section \ref{sec:bias-var} we discuss methods for generating this structure and their tradeoff and their tradeoffss.
The following figure shows an example label tree for a classification problem with 9 classes.

\begin{center}
\begin{tikzpicture}
    [ level distance=1.5cm
    , level 1/.style={sibling distance=3cm}
    , level 2/.style={sibling distance=1.5cm}
    ]
\node {\textit{pseudo1}}
    child {node {\textit{pseudo2}}
      child {node {\texttt{boat}}}
      child {node {\texttt{plane}}}
      child {node {\textit{pseudo3}}
        child {node {\texttt{car}}}
        child {node {\texttt{truck}}}
        child {node {\texttt{van}}}
        child {node {\texttt{motorcycle}}}
      }
      }
    child {node {\texttt{building}}}
    child {node {\textit{pseudo4}}
      child {node {\texttt{cat}}}
      child {node {\texttt{dog}}}
    };
\end{tikzpicture}
\end{center}

We parameterize the weight vectors from the label tree as follows.
For each leaf node $i$ (equivalently for each class $i$),
we let the sequence $P_i$ denote the path from the leaf node to the root.
For example, using the label tree above,
we have
\begin{equation}
    P_{\texttt{truck}} = (\texttt{truck}, \textit{pseudo3}, \textit{pseudo2}, \textit{pseudo1} ) \qquad\text{and}\qquad
    P_{\texttt{cat}} = (\texttt{cat}, \textit{pseudo4}, \textit{pseudo1} )
    .
\end{equation}
Next, we associate a vector $\vv_i$ to each leaf and internal node $i$ in the tree structure.
In particular, there is a there is a vector $\vv_i$ for each class and for each pseudo-class.
Finally, we reparameterize the weights for the softmax cross entropy loss as the sum of the vectors along the paths.
Specifically, we set the weight vector for class $j$ to be 
\begin{equation}
    \label{eq:covertree-reparam}
    \w_j = \sum_{k\in P_j} \vv_k
    .
\end{equation}
Substituting \eqref{eq:covertree-reparam} into \eqref{eq:xentropy} gives us the final cover tree loss
\begin{equation}
    \ell(V;(\x,y)) = - \log \frac {\exp(-\sum_{k\in P_y}\trans\vv_k \x)}{\sum_{j=1}^c \exp(- \sum_{k\in P_j}\trans\vv_k \x)}
\end{equation}
where $V=(\vv_1,...,\vv_k)$,
and our goal is to minimize $\ell$ with respect to $V$.
Note that this loss can easily be computed in libraries like pytorch or tensorflow with the standard cross entropy loss functions.
The only change needed is to define the loss in terms of the sum of vectors $V$ instead of a variable tensor $W$.

@@

    Next, we show that
    \begin{equation}
        \left(\frac q 2\right)^r
        =
        \left(\frac q 2\right)^{\log_q n}
        =
        \frac{n}{2^{\log_q n}}
        =
        \frac{n}{2^{{\log_2 n}/{\log_2 q}}}
        =
        \frac{n}{n^{1/\log_2 q}}
        =
        n^{\left(1 - {1}/{\log_2 q}\right)}
        .
        \label{eq:lem:technical2}
    \end{equation}
    Substituting \eqref{eq:lem:technical2} into \eqref{eq:lem:lFV:3} gives the stated result.
\end{proof}

@@

\begin{theorem}
\label{theorem:xentropy}
% Assume that for all $i\in[n]$, $\ltwo{x_i} \le \rho$,
%     and that for each class $i\in[c]$, $\ltwo{\w_i}\le B$.
\fixme{Add assumptions.}
Assume that for all $i\in[n]$, $\ltwo{x_i} \le \rho$,
    and that for each class $i\in[c]$, $\ltwo{\w_i}\le \frac{B}{\sqrt k}$.
Then the parameter vector $\bar W$ estimated from one epoch of SGD satisfies
\begin{equation}
    \E L_D(\bar W) - L_D(W^*) \le \frac {B\rho}{\sqrt n}
    ,
\end{equation}
which is independent of the number of classes $c$.
\end{theorem}
%Let $\mathcal L$ be a metric space over the set of labels,
%and $\ddim$ be the doubling dimension of $\mathcal L$.
%Then we build a ``cover tree'' from the class labels that respects the metric structure.
%This cover tree will have exactly $c$ leaf nodes, one for each class label,
%and $O(c)$ internal nodes.
%Why does this small change work?
%Intuitively, classes that have a small distance from each other should have similar weight vectors.

@@

With our problem defined in this way, it is clear that
\begin{equation}
    \argmin_{W} \tfrac 1 n \ell(W; \x_i, y_i)
\end{equation}
converges to $W^*$ using the standard cross entropy loss and both the matrix factorization and cover tree reparameterizations.
\fixme{You should prove this for yourself.}
Figure \ref{fig:synthetic-results} shows that the cover tree loss converges at a faster rate.

\begin{figure}[h]
    \centering
    \subfigure{
        \label{Fig.sub.1}
        \includegraphics[width=0.3\textwidth]{fig/images/loss_vs_n.png}}
    \subfigure{
        \label{Fig.sub.2}
        \includegraphics[width=0.3\textwidth]{fig/images/error_vs_n.png}}
    \subfigure{
        \label{Fig.sub.3}
        \includegraphics[width=0.3\textwidth]{fig/images/accuracy_vs_n.png}}
    \caption{Loss value, parameter matrix error, accuracy versus number of data points(Dimension: 64, Randomness: 1.0, Number of Classes: 10)}
    \label{Fig.1}

    \subfigure{
        \label{Fig.sub.1}
        \includegraphics[width=0.3\textwidth]{fig/images/loss_vs_d.png}}
    \subfigure{
        \label{Fig.sub.2}
        \includegraphics[width=0.3\textwidth]{fig/images/error_vs_d.png}}
    \subfigure{
        \label{Fig.sub.3}
        \includegraphics[width=0.3\textwidth]{fig/images/accuracy_vs_d.png}}
    \caption{Loss value, parameter matrix error, accuracy versus dimension of parameter matrix(Number of Data Points: 100, Randomness: 1.0, Number of Classes: 10)}
    \label{Fig.2}

    \subfigure{
        \label{Fig.sub.1}
        \includegraphics[width=0.3\textwidth]{fig/images/loss_vs_sigma.png}}
    \subfigure{
        \label{Fig.sub.2}
        \includegraphics[width=0.3\textwidth]{fig/images/error_vs_sigma.png}}
    \subfigure{
        \label{Fig.sub.3}
        \includegraphics[width=0.3\textwidth]{fig/images/accuracy_vs_sigma.png}}
    \caption{Loss value, parameter matrix error, accuracy versus randomness(Number of Data Points: 100, Dimension: 64, Number of classes: 10)}
    \label{Fig.3}

    \subfigure{
        \label{Fig.sub.1}
        \includegraphics[width=0.3\textwidth]{fig/images/loss_vs_class.png}}
    \subfigure{
        \label{Fig.sub.2}
        \includegraphics[width=0.3\textwidth]{fig/images/error_vs_class.png}}
    \subfigure{
        \label{Fig.sub.3}
        \includegraphics[width=0.3\textwidth]{fig/images/accuracy_vs_class.png}}
    \caption{Loss value, parameter matrix error, accuracy versus number of classes(Number of Data Points: 100, Dimension: 64, Randomness: 1.0)}
    \label{Fig.4}

    \caption{
        \fixme{We need a plot of loss vs number of data points ($n$), vs dimension ($d$), and vs randomness ($sigma$). These plots should all be log-log plots.  For each of these experiments, you should write to shell scripts: the first to generate the data and the second to plot the data.  These shell scripts should be easy to run in order to easily replicate your experiments.}}
    \label{fig:synthetic-results}
\end{figure}

@@

In multilabel classification, it is common for the classes to be unbalanced.
We would like to show that the cover tree loss works well for unbalanced data.
A good way to do that is to change the distribution that $y_i$ is drawn from,
and a geometric distribution is a likely good alternative.

The covertree loss can work well even when the weight matrix has non-linear intrinsic structure.
We should adapt the $\w*_i$ vectors to sit on a manifold instead of a hyperplane to demonstrate this fact.

\subsection{Pseudo-synthetic data}
In a pseudo-synthetic dataset, we take a non-synthetic dataset and modify it.
\fixme{describe your procedure for generating pseudo-synthetic data and all the experiments that you plan to run}
\subsubsection{Constructing Pseudo-synthetic datasets}
\label{sec:pseu-syn}
In order to compare the performance between classic cross entropy loss function and cover tree loss function, we need datasets which the classes have clear similarities.
Besides, we would like to construct pseudo-synthetic datasets which we can control the similarities.
The idea is we can generate the new classes from the original classes of datasets.
It means only the labels are synthetic and the features are real.
We get the similarities of new classes from probability distribution and control the complexity of similarities by changing probability distribution.

For non-synthetic datasets that do not have specific similarities, we would build pseudo-synthetic datasets with similarities manually by following procedures. 
\begin{itemize}
  \item
  For $o$ original classes, we want to generate a matrix $M : \R^{c\times o}$ for c new classes which each row represents the probability distribution of a new class. 
  Each new class is uniformly generated from $k$ original classes.
  So $M$ is as below:
\begin{align}
    r &= (j+l) \bmod c \\
    M_{i+r,l} &= \frac{o}{ck}  
\end{align}
where $i \in [c]$ $l \in [o]$ $j \in [k]$

  \item
  We generate new classes by the matrix $M$.
  To be specific, each column of $M$ represents the probability distribution of each original class.
  For each original class, we generate new classes by corresponding column of $M$.
  For example, we have $100$ data points with original class $1$ and $M_{1,1}=0.5$ $M_{2,1}=0.5$, all other elements of the first column of $M$ are $0$.
  From the first column of $M$, we generate $50$ data points with new class $1$ and $50$ data points with new class $2$.
  If so, we can generate a pseudo-synthetic datasets with new classes $c$.

  For each data point $i\in[n]$, 
  $y$ are new classes and $Y$ are original classes,
  we sample the data point according to the following rules:
\begin{align}
    y_i &\sim \text{Uniform}([r_1, r_2,...,r_j,...])
\end{align}
where $r_j$ is row number of $M$ and $M_{r_j,Y_i} \ne 0$

  We view each row of $M$ as vector representation of each new class and then get similarities from the vector representations.

  \item
  Next, we are able to construct a cover tree structure from the new classes and take each leaf node represents a new class and replace all parent nodes with pseudo classes. 
\end{itemize}

\subsubsection{Pseudo-synthetic experiments}
As for pseudo-synthetic experiments, we select three datasets: MNIST, CIFAR10 and CIFAR100. 
Table 1 shows the parameters of three datasets. We modify the three non-synthetic datasets to pseudo-synthetic datasets. 
The point is we want to construct pseudo-synthetic datasets with $c=10o$ classes as described in Section~\ref{sec:pseu-syn}.
We would like to run experiments with different complexity of similarities by changing $k$.
To start with $k=1$, the elements of similarity is $0$ or $1$.
Then we take image classification task on pseudo-synthetic datasets.

\begin{table}[h]
    \centering
    \input{fig/datasets.tex}
\end{table}

\subsection{Real world twitter data}
\fixme{Write a description of the experiment.}

>>>>

\documentclass[twoside]{article}

\usepackage{aistats2022}
% If your paper is accepted, change the options for the package
% aistats2022 as follows:
%
%\usepackage[accepted]{aistats2022}
%
% This option will print headings for the title of your paper and
% headings for the authors names, plus a copyright note at the end of
% the first column of the first page.

% If you set papersize explicitly, activate the following three lines:
%\special{papersize = 8.5in, 11in}
%\setlength{\pdfpageheight}{11in}
%\setlength{\pdfpagewidth}{8.5in}

% If you use natbib package, activate the following three lines:
%\usepackage[round]{natbib}
%\renewcommand{\bibname}{References}
%\renewcommand{\bibsection}{\subsubsection*{\bibname}}

% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{natbib}
\usepackage[inline]{enumitem}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{subfigure}

@@

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
%\runningtitle{I use this title instead because the last one was very long}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}

\twocolumn[

\aistatstitle{The Cover Tree Loss}

\aistatsauthor{ Yujie Wang \And Michael Izbicki  }

\aistatsaddress{ Claremont Graduate University \And  Claremont McKenna College } ]

\begin{abstract}
Softmax Cross Entropy loss function is widely used in neural network multi-class classification tasks. 
It treats every class equally while classes often have an internal structure. 
We introduce cover tree loss function which takes advantage of this internal class structure.
Compare to softmax cross entropy which has generalization error $O(\sqrt{c/n})$, the generalization error of cover tree loss is $O(\sqrt{1/n})$ which is independent of the number of classes. 
We provide the theoretical justification and conduct extensive experiments on fully synthetic data and real world data. 
The cover tree loss outperforms than softmax cross entropy loss function on metrics.  
\end{abstract}

\section{Introduction}
When people do a standard multi-class classification task, the standard loss function is softmax cross entropy loss function.
The softmax cross entropy loss function

@@

The main theoretical result for the cross entropy loss is that for a dataset with $n$ points,
the generalization error is $O(\sqrt{c/n})$.
There are many theorems that establish this result in different contexts.
The following is an example that relies on properties of SGD.
\begin{theorem}
\label{theorem:xentropy}
Assume that for all $i\in[n]$, $\ltwo{x_i} \le \rho$,
    and that for each class $i\in[c]$, $\ltwo{\w_i}\le B$.
Then the parameter vector $\bar W$ estimated from one epoch of SGD satisfies
\begin{equation}
    \E L_D(\bar W) - L_D(W^*) \le \frac {\sqrt cB\rho}{\sqrt n}
\end{equation}
where $W^* = \argmin L_D(W)$.
\end{theorem}
% \begin{proof}
%     \fixme{Write a proof of this result using Theorem 14.12 of Shalev-Shwartz and Ben-David and the properties of the cross entropy loss from the last set of notes.}
%     \begin{align}
%     {\| W \|^2_F}&=\sum_{i=1}^c{\| W_i\|^2}\\
%                 &\le\sum_{i=1}^c{B^2}\\
%                 &\le cB^2\\
%   \| W \|_F  &\le \sqrt c B 
%     \end{align}
% Then we get:
%     \begin{equation}
%     l(\bar W)-l(W^*)\le\frac {\sqrt cB\rho}{\sqrt n}
%     \end{equation}
%   Take expectation of both side of equation(3):
%   \begin{equation}
%     \E L_D(\bar W) - L_D(W^*) \le \frac {\sqrt cB\rho}{\sqrt n}
% \end{equation}
% \end{proof}

@@

HSM:
\begin{equation}
    \label{eq:xentropy}
    \ell(W;(\x,y)) = - \log \prod_{k\in P_y}\frac {\exp(-\trans\w_k \x)}{\sum_{j=1}^c \exp(-\trans \w_j \x)}
\end{equation}
where $P_y$ is path from root to node of Huffman Tree

% \fixme{Read the papers mentioned above.  Verify that you agree with the things I wrote.  Then, write a detailed mathematical description of the HSM loss and contrast it with the cover tree loss.}

% \fixme{Read 5-10 additional papers about the hierarchical softmax.  For each paper, write a paragraph describing the paper's contributions and how it relates to the cover tree loss.  For the final paper we will reference these papers.  We will have to trim down the notes that you write for this task, but it's useful when writing the paper to have detailed notes about references.}

@@

% \fixme{Write a detailed description of the sim loss}

Simloss \cite{Kobs2020SimLossCS} is similarity based loss function which incorporates class similarities to support the training of neural network classifiers.
Cross entropy loss assumes only one class is correct and punishes every misclassification, while Simloss indicates classes have similarities and classifying a target as a similar class is better.
When calculating the loss function, instead of only taking the probability vector at the target index Simloss also adds the product of similarity between target and other classes and probability vector at the corresponding class. 
Non zero values of similarity matrix lead to smaller losses when the network gives a high score to classes similar to the correct one. 
It would allow the classifier to make less severe mistakes as it learns to predict similar classes.
\begin{align}
    L_{Simloss}&=-\frac{1}{N}\log(\sum_{c=1}^{C}S_{y_i,c}\cdot p_i[c])\\
    % L_{Simloss}=-\frac{1}{N}\log(\sum_{c=1}^{C}S_{y_i,c}\cdot\frac {\exp(-\trans\w_{y_i} \x)}{\sum_{j=1}^C \exp(-\trans \w_j \x)})
    p_i&=\frac {\exp(-\trans\w_{y_i} \x)}{\sum_{j=1}^C \exp(-\trans \w_j \x)}
\end{align}

@@

This factorization decreases the statistical complexity of the problem from $O(\sqrt{c/n})$ to $O(\sqrt{e/n})$.
Recall that $e$ is a hyperparameter that is selected to be much less than $c$ and should represent the "intrinsic" dimensionality of the weight matrix.
The following theorem proves this result formally.
\begin{theorem}
Assume that for all $i\in[n]$, $\ltwo{x_i} \le \rho$,
    and that for each class $i\in[c]$, $\ltwo{\vv_i}\le 1$ and $\ltwo{\uu_i}\le B$.
Then the parameter vector $\bar W = V \bar U$ estimated from one epoch of SGD satisfies
\begin{equation}
    \E L_D(\bar W) - L_D(W^*) \le \frac {\sqrt eB\rho}{\sqrt n}
\end{equation}
where $W^* = \argmin L_D(W)$.
\end{theorem}
% \begin{proof}
%     \fixme{Write a proof of this result using Theorem 14.12 of Shalev-Shwartz and Ben-David.}
%     \begin{align}
%     {\| U \|^2_F}&=\sum_{i=1}^e{\| U_i\|^2}\\
%                 &\le\sum_{i=1}^e{B^2}\\
%                 &\le eB^2\\
%   \| U \|_F  &\le \sqrt e B  \\
%   \| W \|_F  &\le \sqrt e B
%     \end{align}
%     Then we get:
%     \begin{equation}
%     l(\bar W)-l(W^*)\le\frac {\sqrt eB\rho}{\sqrt n}
%     \end{equation}
%   Take expectation of both side of equation(8):
%   \begin{equation}
%     \E L_D(\bar W) - L_D(W^*) \le \frac {\sqrt eB\rho}{\sqrt n}
% \end{equation}
% \end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Tree Loss}

Like the weight matrix factorization, the tree loss works by reparameterizing the weight matrix for the softmax cross entropy loss.
The reparameterization is defined according to a \emph{label tree} structure that captures the similarity between labels.
Every class is represented by a leaf node in the label tree,
and internal nodes represent \emph{pseudo-classes} that group the existing classes into a logical hierarchy.
At this point, we assume that the label tree structure is given a priori, and in Section \ref{sec:bias-var} we discuss methods for generating this structure and their tradeoffs.
The following figure shows an example label tree for a classification problem with 9 classes.

\begin{center}
\small
\begin{tikzpicture}
    [ level distance=1.5cm
    , level 1/.style={sibling distance=3cm}
    , level 2/.style={sibling distance=1.5cm}
    ]
\node {\textit{pseudo1}}
    child {node {\textit{pseudo2}}
      child {node {\texttt{boat}}}
      child {node {\texttt{plane}}}
      child {node {\textit{pseudo3}}
        child {node {\texttt{car}}}
        child {node {\texttt{truck}}}
        child {node {\texttt{van}}}
        child {node {\texttt{motorcycle}}}
      }
      }
    child {node {\texttt{building}}}
    child {node {\textit{pseudo4}}
      child {node {\texttt{cat}}}
      child {node {\texttt{dog}}}
    };
\end{tikzpicture}
\end{center}

We parameterize the weight vectors from the label tree as follows.
For each leaf node $i$ (equivalently for each class $i$),
we let the sequence $P_i$ denote the path from the leaf node to the root.
For example, using the label tree above,
we have

\begin{equation}
\begin{split}
    P_{\texttt{truck}} &= (\texttt{truck}, \textit{pseudo3}, \textit{pseudo2}, \textit{pseudo1} ) \\
    P_{\texttt{cat}} &= (\texttt{cat}, \textit{pseudo4}, \textit{pseudo1} )
\end{split}
\end{equation}

% \end{equation}

Next, we associate a vector $\vv_i$ to each leaf and internal node $i$ in the tree structure.
In particular, there is a there is a vector $\vv_i$ for each class and for each pseudo-class.
Finally, we reparameterize the weights for the softmax cross entropy loss as the sum of the vectors along the paths.
Specifically, we set the weight vector for class $j$ to be 
\begin{equation}
    \label{eq:covertree-reparam}
    \w_j = \sum_{k\in P_j} \vv_k
    .
\end{equation}
Substituting \eqref{eq:covertree-reparam} into \eqref{eq:xentropy} gives us the final cover tree loss
\begin{equation}
    \ell(V;(\x,y)) = - \log \frac {\exp(-\sum_{k\in P_y}\trans\vv_k \x)}{\sum_{j=1}^c \exp(- \sum_{k\in P_j}\trans\vv_k \x)}
\end{equation}
where $V=(\vv_1,...,\vv_k)$,
and our goal is to minimize $\ell$ with respect to $V$.
Note that this loss can easily be computed in libraries like pytorch or tensorflow with the standard cross entropy loss functions.
The only change needed is to define the loss in terms of the sum of vectors $V$ instead of a variable tensor $W$.

@@

    Next, we show that
    \begin{equation}
    \begin{split}
        \left(\frac q 2\right)^r
        &=
        \left(\frac q 2\right)^{\log_q n}\\
        &=
        \frac{n}{2^{\log_q n}}\\
        &=
        \frac{n}{2^{{\log_2 n}/{\log_2 q}}}\\
        &=
        \frac{n}{n^{1/\log_2 q}}\\
        &=
        n^{\left(1 - {1}/{\log_2 q}\right)}
        .
        \label{eq:lem:technical2}
    \end{split}
    \end{equation}
    Substituting \eqref{eq:lem:technical2} into \eqref{eq:lem:lFV:3} gives the stated result.
\end{proof}

@@

\begin{theorem}
\label{theorem:xentropy}
% Assume that for all $i\in[n]$, $\ltwo{x_i} \le \rho$,
%     and that for each class $i\in[c]$, $\ltwo{\w_i}\le B$.
% \fixme{Add assumptions.}
Assume that for all $i\in[n]$, $\ltwo{x_i} \le \rho$,
    and that for each class $i\in[c]$, $\ltwo{\w_i}\le \frac{B}{2^{\level k}}$.
Then the parameter vector $\bar W$ estimated from one epoch of SGD satisfies
\begin{equation}
    \E L_D(\bar W) - L_D(W^*) \le \frac {B\rho}{\sqrt n}
    ,
\end{equation}
which is independent of the number of classes $c$.
\end{theorem}
%Let $\mathcal L$ be a metric space over the set of labels,
%and $\ddim$ be the doubling dimension of $\mathcal L$.
%Then we build a ``cover tree'' from the class labels that respects the metric structure.
%This cover tree will have exactly $c$ leaf nodes, one for each class label,
%and $O(c)$ internal nodes.
%Why does this small change work?
%Intuitively, classes that have a small distance from each other should have similar weight vectors.

@@

With our problem defined in this way, it is clear that
\begin{equation}
    \argmin_{W} \tfrac 1 n \ell(W; \x_i, y_i)
\end{equation}
converges to $W^*$ using the standard cross entropy loss and both the matrix factorization and cover tree reparameterizations.

\begin{figure}[h]
\vspace{.3in}
\centering
\includegraphics[width=0.3\textwidth]{fig/images/accuracy_vs_n.png}
\vspace{.3in}
\caption{Accuracy versus number of data points($d=64, \sigma=1.0, c=10$)}
\label{Fig.1}
\end{figure}

\begin{figure}[h]
\vspace{.3in}
\centering
\includegraphics[width=0.3\textwidth]{fig/images/accuracy_vs_d.png}
\vspace{.3in}
\caption{Accuracy versus dimension of parameter matrix($n=100, \sigma=1.0, c=10$)}
\label{Fig.2}
\end{figure}

\begin{figure}[h]
\vspace{.3in}
\centering
\includegraphics[width=0.3\textwidth]{fig/images/accuracy_vs_sigma.png}
\vspace{.3in}
\caption{Accuracy versus randomness($n=100, d=64, c=10$)}
\label{Fig.3}
\end{figure}

\begin{figure}[h]
\vspace{.3in}
\centering
\includegraphics[width=0.3\textwidth]{fig/images/accuracy_vs_class.png}
\vspace{.3in}
\caption{Accuracy versus number of classes($n=100, d=64, \sigma=1.0$)}
\label{Fig.4}
\end{figure}

% \subsubsection{Analysis of Synthetic Experiment results}
% From \ref{Fig.1}, as number of data points increases loss value and parameter matrix error decrease while accuracy increases. 
% We compare the performance of cover tree loss, softmax cross entropy loss and simloss, it is clear that cover tree loss converges faster than the other two loss functions.
% Besides, cover tree loss achieves lower loss value and parameter matrix error over number of data points.

% As shown in \ref{Fig.2}, loss value decreases and accuracy increases over dimension. 
% Similar to \ref{Fig.1}, cover tree loss converges faster and achieve lower loss value.
% We take the line of parameter matrix error of cover tree loss as a benchmark and draw the difference of parameter matrix error between cover tree loss and the other two loss functions.
% The middle sub-figure indicates that cover tree loss has lower parameter matrix error with dimension than cross entropy and simloss.

% According to \ref{Fig.3}, if we increase the randomness and fix the other parameters, loss value increases and accuracy decreases.
% Cover tree loss performs similarly with cross entropy loss and simloss in terms of loss value and accuracy.
% The middle sub-figure implies that parameter matrix error of cover tree loss decreases even though the randomness increases.

% \ref{Fig.4} indicates that, as number of classes increases, the loss value and parameter matrix error also increases and accuracy certainly decrease.
% From middle sub-figure, the parameter matrix error of cover tree loss is lower than the other two loss functions.
% The right sub-figure shows that the accuracy of cover tree loss decreases slower and can keep higher value than the other two loss functions.

@@

In multi-label classification, it is common for the classes to be unbalanced.
We would like to show that the cover tree loss works well for unbalanced data.
A good way to do that is to change the distribution that $y_i$ is drawn from,
and a geometric distribution is a likely good alternative.

The cover tree loss can work well even when the weight matrix has non-linear intrinsic structure.
We should adapt the $\w*_i$ vectors to sit on a manifold instead of a hyperplane to demonstrate this fact.

\subsection{Real world data}
% \fixme{Write a description of the experiment.}
In this experiment, we will compare cover tree loss with softmax corss entropy and simloss on some real world data.
We conduct image classification tasks on CIFAR-100 dataset and ImageNet dataset and emoji prediction task on twitter dataset.
Our goal is to investigate the effects of changing the loss function on our example experiments.
We do not focus on if we have state-of-the-art performance on specific models, but rather on the evaluation of cover tree loss as a general loss function which can be applied on various tasks.
These three tasks and datasets are typical examples for using softmax cross entropy loss.

\subsubsection{CIFAR-100 Data}
We use CIFAR-100 dataset which has 100 classes for image classification.
We also get word embeddings from a word2vec model pretrained on Google News \cite{Mikolov2013EfficientEO} of class names and construct cover tree structure and calculate similarity between classes from word embeddings.
If the class name include more than one word, we take a sum of the embedding of each word.

To evaluate our cover tree loss, we rely on two metrics: accuracy (Acc.) which focus on correct predictions and similarity accuracy (SA) which focus on the similarity of predicted and target class.
We take ResNet20 model \cite{He2016DeepRL} and SGD optimizer with a learning rate 0.1 and a batch size of 1021. 
We stop early to get the highest accuracy on validation set.

\begin{table}[h]
\small
\caption{Validation and test results with early stopping on CIFAR-100 dataset. Best values are written in bold.} \label{cifar100}
\begin{center}
\begin{tabular}{@{}lcccc@{}}
\toprule
\multicolumn{1}{c}{\multirow{2}{*}{Loss}} & \multicolumn{2}{c}{VALIDATION} & \multicolumn{2}{c}{TEST}       \\ \cmidrule(l){2-5} 
\multicolumn{1}{c}{}                      & Acc.       & SA            & Acc.      & SA             \\ \cmidrule(r){1-1}
Cover Tree Loss                           & \textbf{53.44} & \textbf{66.7} & \textbf{53.5} & \textbf{65.89} \\
Cross Entropy                             & 52.08          & 65.46         & 51.87         & 65.58          \\
SimLoss                                   & 52.50           & 65.76         & 52.43         & 65.82          \\ \bottomrule
\end{tabular}
\end{center}
\end{table}

\subsubsection{ImageNet Data}
ImageNet dataset \cite{Russakovsky2015ImageNetLS} has 1000 classes and many of the classes are similar and related, we consider the internal class structure has a impact on performance of loss function.
We use class names to get class representations from a pretrained BERT model \cite{Devlin2019BERTPO}.
We take ResNet50 model \cite{He2016DeepRL} and SGD optimizer with a learning rate 0.01.
Our metrics are top 1 accuracy and top 5 accuracy.

\begin{table}[h]
\small
\caption{Validation and test results on ImageNet dataset. Best values are written in bold.} \label{imagenet}
\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
\multicolumn{1}{c}{\multirow{2}{*}{Loss}} & \multicolumn{2}{c}{VALIDATION} \\ \cmidrule(l){2-3} 
\multicolumn{1}{c}{}                      & Top1 Acc.      & Top5 Acc.     \\ \cmidrule(r){1-1}
Cover Tree Loss                           &                &               \\
Cross Entropy                             &                &               \\
SimLoss                                   &                &               \\ \bottomrule
\end{tabular}
\end{center}
\end{table}

\subsubsection{Twitter Data}
